# my little reminder for some aspects of the lattice charts

```{r}
df1<-data.frame(l=rep(letters[1:10],2), 
               v1=abs(rnorm(20))*10, 
               f=c(rep(c("f1"),10),rep(c("f2"),10)))
df1
```

### plain bare bone chart
```{r}
require(lattice)
xyplot(v1~l|f, data=df1)
```

### use groups instread of conditioning | 
```{r}
xyplot(v1~l, groups=f, data=df1)
```

### failed attempt to align the grid lines with x labels
#### to note panel.grid() is not working properly (but just sometime!)

```{r}

xyplot(v1~l, groups=f, data=df1,
                 
       panel=function(...){
         
         panel.grid(v=-1,h=-1,col = "grey", ...)
         
         panel.xyplot(...)
       }
       
       )
       
```
### tweak the alignment of grid lines with x labels by using panel.abline()

```{r}
xyplot(v1~l, groups=f, data=df1,
                 
       panel=function(...){
         
         panel.grid(v=0,h=-1,col = "grey", ...)
         panel.abline(v=1:10,col = "grey", lwd = 1, lty = 2)
         panel.xyplot(...)
       }
       
       )
```

### further chart customization 
- color and size of points (see par.settings and simpleTheme)
- x axis labels (see scale, capital letters now)
- legend (see auto.key)

```{r}
xyplot(v1~l, groups=f, data=df1,
             
       par.settings=simpleTheme(col.points=c("red","green"), cex=1.2, pch=19),       
       scale = list(x = list(at = 1:10,labels = LETTERS[1:10], cex = 0.8)),
       auto.key = list(points=TRUE,columns=2, position="top"),
       
       panel=function(...){
         
         panel.grid(v=0,h=-1,col = "grey", ...)
         panel.abline(v=1:10,col = "grey", lwd = 1, lty = 2)
         panel.xyplot(...)
       }
       
       )

```
### dealing with dates on the x axis

#### define a new dataframe
```{r}
df2 <- data.frame(date = seq(as.Date("2014/01/01"), as.Date("2014/01/31"), by="day"),
                  v1 = runif(31)*10)
```

#### x labels as default
```{r}
xyplot(v1 ~ date, data = df2)
```
#### one label for each original timestamp and aligned grid line
```{r}
mydate<-df2$date
xyplot(v1 ~ date, data = df2,
       
       panel = function(...) {
         panel.grid(h = -1, v = 0, col = "grey", lwd = 1, lty = 1)
         panel.abline(v = df2$date, col = "grey", lwd = 1, lty = 1)
         panel.xyplot(...)
         },
       scale = list(x = list(at = mydate, labels = format(mydate, "%d-%b-%y"), cex = 0.8, rot=90)) 
       )
```

#### customized coarser labels and aligned grid line

```{r}
nlab <- 6 # choose number of labels
mydate2 <- seq(from = min(df2$date), to = max(df2$date), length.out = nlab)

xyplot(v1 ~ date, data = df2,
       
       panel = function(...) {
         panel.grid(h = -1, v = 0, col = "grey", lwd = 1, lty = 1)
         panel.abline(v = df2$date, col = "grey", lwd = 1, lty = 1)
         panel.xyplot(...)
       },
       scale = list(x = list(at = mydate2, labels = format(mydate2, "%d-%b-%y"), cex = 0.8, rot=90)) 
)
```



